# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r load,echo=T}
data<-read.csv("activity.csv",header=T)
data$newdate<-strptime(data$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r totalSteps,echo=T}
total_num<-tapply(data$steps,data$date,sum,na.rm=T)
hist(total_num)

mean_d<-mean(total_num)
median_d<-median(total_num)
```
The mean total number of steps taken per day is `r mean_d`. The median total number of steps taken per day is `r median_d`.

## What is the average daily activity pattern?
```{r,echo=T}
aver_interval<-tapply(data$steps,data$interval,mean,na.rm=T)
plot(as.numeric(names(aver_interval)),aver_interval,type="l",xlab="intervals",ylab="average steps")

x<-names(aver_interval)[which(aver_interval==max(aver_interval))]
```
The `r x` interval contains the maximum number of steps.


## Imputing missing values
```{r,echo=T}
# total number of missing values
num_na<-sum(is.na(data$steps))

sub_data<-subset(data,is.na(data$steps))
# replace NAs with the average value across 5-interval
new_data<-data # create a new dataset
new_data$steps[is.na(new_data$steps)]=aver_interval[as.character(sub_data$interval)]

# Make a histogram of the total number of steps taken each day 
new_total_num<-tapply(new_data$steps,new_data$date,sum)
hist(new_total_num)
new_mean_d<-mean(new_total_num)
new_median_d<-median(new_total_num)

```
Total number of missing values in the dataset is `r num_na`.
After replacing NA values to average value per 5-interval, the new mean is `r new_mean_d` and the new median is `r new_median_d`.The values are different from the ones from first part of the assignment. The values of mean and median improve by replacing the NAs.

## Are there differences in activity patterns between weekdays and weekends?
```{r,echo=T}
Sys.setlocale("LC_TIME", "en_US")
# create a new factor variable in the dataset
wd<-as.character(weekdays(strptime(data$date,"%Y-%m-%d")))
l<-wd=="Saturday"|wd=="Sunday"
wd[l]="weekend"
wd[!l]="weekday"
wd<-factor(wd)
data$wd<-wd
str(data)

#make a plot about the average number of steps with 5-min interval.
data_weekend<-subset(data,wd=="weekend")
data_weekday<-subset(data,wd=="weekday")
weekend_aver_interval<-tapply(data_weekend$steps,data_weekend$interval,mean,na.rm=T)
weekday_aver_interval<-tapply(data_weekday$steps,data_weekday$interval,mean,na.rm=T)

par(mfcol=c(2,1))
par(mar=c(4,4,2,2))
plot(as.numeric(names(weekday_aver_interval)),weekday_aver_interval,type="l",xlab="intervals",ylab="average steps",main="weekend")
plot(as.numeric(names(weekend_aver_interval)),weekend_aver_interval,type="l",xlab="intervals",ylab="average steps",main="weekday")

```
